KidsAIR: merge exposure datasets

### 2020-04-30 exposure_models.Rmd
ambient temp/pm and exposure model udpates

---
title: "KidsAIR: exposure model"
author: "Ethan Walker"
date: "Started 19 March 2020, Updated 16 April 2020"
output: pdf_document
---

@@

input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
input_file <- c("exposure_analysis_data_long_new.rds")
output_file <- c("exposure_analysis_data_long_new.rds")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
```

@@

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>%
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(treatment == "Placebo") %>% 
  #filter(winter_id == 1) %>% 
  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
  group_by(area, home, home_winter_id) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  mutate_at(vars(starts_with("home_act")), factor) %>%
  ungroup() %>% 
  distinct(area, home,  home_winter_id, .keep_all = TRUE)

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  group_by(home) %>% 
  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
                                   lead(residents_smoke), residents_smoke)) %>% 
  ungroup() %>% 
  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) 
```

@@

# Select analysis vars
selected_vars <- analysis_data %>% 
  select(area, pm_mean_sampling_period, income_3level, residents_smoke, 
         stove_age_3level, home_windows_2level,
         chimney_clean_3level, wood_collect_2level,
         home_act_door, home_act_sweep, home_act_windows, 
         sums_events_sampling_period, mean_temp_sampling_period_5, 
         stove_grade_3level, moisture_closest_2level) %>% 
  mutate(home_act_door = factor(home_act_door, levels = c(0,1,2,3,4,5,6)),
         home_act_sweep = factor(home_act_sweep, levels = c(0,1,2,3,4,5,6)),
         home_act_windows = factor(home_act_windows, levels = c(0,1,2,3,4,5,6)),
         area = factor(area, levels = c("AK", "NN", "WMT")),
         log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
  select(-pm_mean_sampling_period) %>% 
  filter(!is.na(income_3level) & !is.na(home_windows_2level))

@@

# Impute missing data; m = 20 datasets (check % missing data and match)
imputed_data <- mice(selected_vars, m=5, maxit = 5,
                     method = c("polyreg", "polr", "logreg", "polr",
                                "logreg", "polr", "logreg", "polr", "polr",
                                "polr", "pmm", "pmm", "polr", "logreg", "pmm"))

>>>>

---
title: "KidsAIR: exposure model"
author: "Ethan Walker"
date: "Started 19 March 2020, Updated 30 April 2020"
output: pdf_document
---

@@

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
```

@@

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>%
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         amb_pm_24hr:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(treatment == "Placebo") %>% 
  #filter(winter_id == 1) %>% 
  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
  group_by(area, home, home_winter_id) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home,  home_winter_id, .keep_all = TRUE)

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  group_by(home) %>% 
  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
                                   lead(residents_smoke), residents_smoke)) %>% 
  ungroup() %>% 
  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
                                 labels = c("<3", "3+"),
                                 right = FALSE))
```

@@

# Select analysis vars
selected_vars <- analysis_data %>% 
  select(area, pm_mean_sampling_period, income_3level, 
         stove_age_3level, home_windows_2level,
         chimney_clean_3level, 
         home_act_door_2level, home_act_windows_2level, home_act_smoke_2level,
         sums_events_sampling_period, mean_temp_sampling_period_5, 
         amb_pm_sampling_period_5, stove_grade_3level, moisture_closest_2level) %>% 
  mutate(area = factor(area, levels = c("AK", "NN", "WMT")),
         log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
  select(-pm_mean_sampling_period) %>% 
  filter(!is.na(income_3level) & !is.na(home_windows_2level)) %>% 
  filter(area == "WMT") %>% 
  select(-area)

@@

# Impute missing data; m = 20 datasets (check % missing data and match)
imputed_data <- mice(selected_vars, m=5, maxit = 5,
                     method = c("polr", "polr", "logreg", "polr", 
                                "logreg", "logreg", "logreg", "pmm", "pmm",
                                "pmm", "polr", "logreg", "pmm"))

### 2020-04-30 merge_datasets.Rmd
ambient temp/pm and exposure model udpates

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date) %>% 
  distinct(zip, .keep_all = TRUE)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  distinct(zip, .keep_all = TRUE)

wmt_zipcodes <- read_csv(paste0(output_path, "Input/WMT/wmt_zipcodes.csv")) %>% 
  rename_all(tolower)

nn_zipcodes <- read_csv(paste0(output_path, "Input/NN/nn_zipcodes.csv")) %>% 
  rename_all(tolower)

ak_zipcodes <- read_csv(paste0(output_path, "Input/AK/ak_zipcodes.csv")) %>% 
  rename_all(tolower)

kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
  mutate(home = gsub(home, pattern = "_", replacement = ""),
         home = gsub(home, pattern = " ", replacement = "")) %>% 
  left_join(kids_amb_pm_data, by = "zip") %>% 
  select(home, zip, lat_pm, long_pm) %>% 
  left_join(kids_amb_temp_data, by = "zip") %>% 
  select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date) %>% 
  select(-lat_zip, -long_zip, -lat_pm, -long_pm)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date)) %>% 
  select(-lat_zip, -long_zip, -lat_temp, -long_temp)

exposure_analysis_data_long_new <- exposure_analysis_data_long %>%
  # fix incorrect data after checking paper logs
  mutate(total_residents = if_else(home == "WMA438", "6", total_residents),
         total_residents = if_else(home == "WMA480" & winter_id == 1, "13", total_residents),
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         income = as.character(income),
         income = if_else(home == "WMA480" & winter_id == 1, "$40,000 to $49,999", income),
         income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+"))) %>% 
  ungroup() %>% 
  left_join(kids_zipcodes, by = "home") %>% 
  left_join(kids_amb_pm_data, by = c("area", "zip", "sample_date")) %>% 
  left_join(kids_amb_temp_data, by = c("area", "zip", "sample_date")) %>%
  arrange(area, home, home_winter_id, sample_date)

>>>>

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date))

exposure_analysis_data_long_new <- exposure_analysis_data_long %>%
  # fix incorrect data after checking paper logs
  mutate(total_residents = if_else(home == "WMA438", "6", total_residents),
         total_residents = if_else(home == "WMA480" & winter_id == 1, "13", total_residents),
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         income = as.character(income),
         income = if_else(home == "WMA480" & winter_id == 1, "$40,000 to $49,999", income),
         income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+"))) %>% 
  ungroup() %>% 
  left_join(kids_amb_pm_data, by = c("area", "home", "sample_date")) %>% 
  select(-zip, -lat_home, -long_home) %>% 
  left_join(kids_amb_temp_data, by = c("area", "home", "sample_date")) %>%
  arrange(area, home, home_winter_id, sample_date)
